\subsection{Promise 2: No dangling pointers}

%%%

\begin{frame}{Promise 2: No dangling pointers}

\begin{itemize}
	\item Rust programs never try to access a heap-allocated value after it has
		been freed.
	\item By default, no garbage collection or reference counting involved!
	\item Everything is enforced at compile-time.
\end{itemize}

\end{frame}

%%%

\begin{frame}[label={threerules}]{Three Rules}

\begin{block}{Rule 1}
Every value has a single \textbf{owner} at any given time.
\end{block}
\pause
\begin{block}{Rule 2}
You can \textbf{borrow} a reference to a value, so long as the
reference doesn’t outlive the value.
\end{block}
\pause
\begin{block}{Rule 3}
You can only modify (mutate) a value when you have \textbf{exclusive}
(mutable) access to it.
\end{block}

\end{frame}

%%%

\begin{frame}{Ownership}

\begin{itemize}
	\item Variable bindings own their values
	\item A struct owns its fields
	\item An enum owns its values
	\item Every heap-allocated value has a single pointer that owns it
	\item All values are dropped when their owner is dropped
\end{itemize}

\end{frame}

%%%

\begin{frame}[fragile]{Ownership: Scoping}

If a value goes out of scope, the corresponding memory is automatically freed.

\begin{minted}{rust}
{
    let s: String = "Chuchichästli".to_string();
} // s goes out of scope, memory is freed
\end{minted}

\end{frame}

%%%

\begin{frame}[fragile]{Ownership: Move Semantics}

Ownership is moved by default.

\begin{minted}{rust}
let s: String = "Chuchichästli".to_string();

// t1 takes ownership from s
let t1 = s;

// compile-time error: use of moved value s
let t2 = s;
\end{minted}

\end{frame}

%%%

\begin{frame}[fragile]{Ownership: Opt-in Implicit Copy Semantics}

Types that implement the \texttt{Copy} marker trait (more about traits later)
are copied instead of moved. The stdlib implements \texttt{Copy} for all
primitive types.

\begin{minted}{rust}
let pi = 3.1415926f32;
let foo = pi;
let bar = pi; // This is fine!
\end{minted}

\end{frame}

%%%

\begin{frame}[fragile]{Ownership: Opt-in Explicit Copy Semantics}

If you prefer copies to be explicit, you can implement the \texttt{Clone} trait
instead.

\begin{minted}{rust}
let s = "Chuchichästli".to_string();
let t1 = s.clone();
let t2 = s.clone();
\end{minted}

\end{frame}

%%%

\begin{frame}[fragile]{Ownership: Deriving Copy / Clone}

The compiler can automatically derive implementations of \texttt{Copy} and
\texttt{Clone} for us.

\begin{minted}{rust}
#[derive(Copy, Clone)]
struct Color {
    r: u8,
    g: u8,
    b: u8
}
\end{minted}

\end{frame}

%%%

\begin{frame}[fragile]{Ownership: Function Parameters}

But what about this?

\begin{minted}{rust}
fn print_loud(text: String) { println!("{}!!!!!", text); }
let s = "Hello, Hackers".to_string();
print_loud(s);
println!("{}", s);
\end{minted}

\pause
\sep

\begin{minted}[fontsize=\footnotesize]{text}
error[E0382]: borrow of moved value: `s`
 --> src/main.rs:8:20
  |
6 |     let s = "Hello, hackers".to_string();
  |         - move occurs because `s` has type `std::string::String`,
  |           which does not implement the `Copy` trait
7 |     print_loud(s);
  |                - value moved here
8 |     println!("{}", s);
  |                    ^ value borrowed here after move
\end{minted}

\end{frame}

%%%

\begin{frame}[fragile]{Borrowing}

Instead of moving a value, it can also be borrowed.

\begin{minted}{rust}
fn print_loud(text: &String) { println!("{}!!!!!", text); }
let s = "Hello, Hackers".to_string();
print_loud(&s);
println!("Original value was {}", s);
\end{minted}

Many functions can borrow at the same time, because they cannot modify.

\end{frame}

%%%

\begin{frame}[fragile]{Mutable Borrowing}

If you need exclusive (=write) access, you can use mutable borrows.

\begin{minted}{rust}
fn make_loud(text: &mut String) { text.push_str("!!!!!"); };
let mut s = "Hello, Hackers".to_string();
make_loud(&mut s);
println!("New value is {}", s);
\end{minted}

While borrow a mutable reference to a value, that reference is the only way to
access that value at all.
\end{frame}

%%%

\begin{frame}[fragile]{Borrowing prevents moving}

While borrowed, a move must be prevented. Otherwise you might end up with a
dangling pointer.

\begin{minted}{rust}
let x = String::new();
let borrow = &x;
let y = x;
println!("Borrowed value is {}", borrow);
\end{minted}

\sep

\begin{minted}[fontsize=\footnotesize]{text}
error[E0505]: cannot move out of `x` because it is borrowed
 --> src/main.rs:4:13
  |
3 |     let borrow = &x;
  |                  -- borrow of `x` occurs here
4 |     let y = x;
  |             ^ move out of `x` occurs here
5 |     println!("Borrowed value is {}", borrow);
  |                                      ------ borrow later used here
\end{minted}

\end{frame}

%%%

\begin{frame}[fragile]{Lifetimes: A contrived example}

What's the problem here?

\begin{minted}{rust}
let borrow: &String;
{
    let x = String::new();
    borrow = &x;
}
println!("Borrowed val is {}", borrow);
\end{minted}
\end{frame}

%%%

\begin{frame}[fragile]{Lifetimes: A contrived example}

The lifetime of the borrow is longer than the lifetime of `x`.

\begin{minted}[fontsize=\footnotesize]{text}
error[E0597]: `x` does not live long enough
 --> src/main.rs:5:18
  |
5 |         borrow = &x;
  |                  ^^ borrowed value does not live long enough
6 |     }
  |     - `x` dropped here while still borrowed
7 |     println!("Borrowed val is {}", borrow);
  |                                    ------ borrow later used here
\end{minted}

Using lifetime checking, the compiler guarantees that there are no dangling
pointers.

\end{frame}
